---
title: "EDA"
author: Tiffany Xi
date: 2018-10-04
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```


##Data import

```{r data_import, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, date, month, everything())
```

##Start grouping

tibble still the same 
one extra line like:# Groups: name, month [36]/ # Groups: month [12]/ # Groups: name [3]
layer sub-dataset  
remove groups using ungroup()

```{r}
weather_df %>% 
  group_by(month)
```

```{r}
weather_df %>% 
  group_by(name)
```

```{r}
weather_df %>% 
  group_by(name, month)
```

n() means the number of elements
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n())
```


```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(number = n())
```


```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(number = n())
```

count--handy--use count() in place of group_by() and summarize()
```{r}
weather_df %>%
  count(name, month)
```


```{r}
weather_df %>%
  group_by(month) %>%
  summarize(n_obs = n(),
            n_days = n_distinct(date))
```

##Grouped summaries

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n(),
            num_na = sum(is.na(tmax)),
            mean_tmax = mean(tmax))
```

#different from mutate----repeat the number to fill in the new column

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n(),
            mean_tmax = mean(tmax, na.rm = TRUE))
#na.rm = TRUE--remove mising value--default set to FALSE
```

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n(),
            mean_tmax = mean(tmax, na.rm = TRUE),
            mean_prec = mean(prcp, na.rm = TRUE),
            median_tmax = median(tmax, na.rm = TRUE),
            sd_tmin = sd(tmin, na.rm = TRUE))
```

This is still a tibble!

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n(),
            median_tmax = median(tmax, na.rm = TRUE),
            sd_tmin = sd(tmin, na.rm = TRUE),
            mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point()
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")
```

plot two tables at once

```{r}
month_df = weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n(),
            median_tmax = median(tmax, na.rm = TRUE),
            sd_tmin = sd(tmin, na.rm = TRUE),
            mean_tmax = mean(tmax, na.rm = TRUE))

  ggplot(weather_df, aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  geom_line(data = month_df, aes(x = month, y = mean_tmax, color = name))
```


Tidy is not always best for human-readable

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
  spread(key = name, value = mean_tmax) %>% 
  knitr::kable(digits = 2)
```

```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>% 
  spread(key = month, value = mean_tmax) %>% 
  knitr::kable(digits = 1)
```

##Grouped mutate

```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(mean_tmax = mean(tmax, na.rm = TRUE))
```

More interesting example

```{r}
weather_df %>%
  group_by(name) %>%
  mutate(centered_tmax = tmax - mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```

##Window functions

Use ranks

Coldest day in every place
```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(tmax_ranking = min_rank(tmax)) %>% 
  filter(tmax_ranking < 2)
```

Keep the three days with the highest max temperature
```{r}
weather_df %>%
  group_by(name, month) %>%
  filter(min_rank(desc(tmax)) < 4)
```

Use lags--offset--used to compare an observation to it’s previous value
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(lag = lag(tmax))
```

##How neat is it?

How much do they change from day to day? quantify the day-by-day variability in max temperature
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax)) %>%
  summarize(temp_change_sd = sd(temp_change, na.rm = TRUE),
            temp_change_max = max(temp_change, na.rm = TRUE))
```


##PUSLE data
```{r}
pulse_data = haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  gather(key = visit, value = bdi, bdi_score_bl:bdi_score_12m) %>%
  separate(visit, into = c("remove_1", "remove_2", "visit"), sep = "_") %>%
  select(id, visit, everything(), -starts_with("remove")) %>%
  mutate(visit = replace(visit, visit == "bl", "00m"),
         visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

pulse_data %>% 
  group_by(visit) %>% 
  summarize(mean_bdi = mean(bdi, na.rm = TRUE),
            median_bdi = median(bdi, na.rm = TRUE)) %>% 
  knitr::kable(digits = 3)
```

##FAS data

```{r}
pup_data = read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(wt_gain = gd18_weight - gd0_weight,
         day_of_tx = as.numeric(day_of_tx))

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>% 
  group_by(dose, day_of_tx) %>% 
  summarize(mean_pivot = mean(pd_pivot, na.rm = TRUE)) %>% 
  filter(!is.na(dose)) %>% 
  spread(key = dose, value = mean_pivot) %>% 
  knitr::kable(digits = 3)
```

