---
title: "Visualization Pt 1"
author: Tiffany Xi
date: 2018-09-27
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
library(ggridges)
```

```{r data_import, cache = TRUE}
library(rnoaa)

weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df %>% View
```

## Start a plot

Blank plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```

Scatterplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```


```{r}
weather_df %>%
  filter(name == "CentralPark_NY") %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

```{r}
weather_sp = 
  ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

```{r}
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()
```

## More plot options

Add an aesthetic

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```

Add a geom
add a smooth line (global function)

se = FALSE---get rid of confidence interval
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) + 
  geom_smooth(se = FALSE)
```

Add some options ...

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .4) +
  geom_smooth(se = FALSE)
```
alpha doesn't work for geom_smooth


color = name is a global map
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE)
```
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .4) +
  geom_smooth(aes(color = name), se = FALSE)
```
same result(less straightforward)

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = prcp)) + 
  geom_point(alpha = .4)
```

Facetting ...

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .4) +
  geom_smooth(se = FALSE) + 
  facet_grid( ~ name)
```


A more interesting plot

Precipitation

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name, size = prcp)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(~ name)
```

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_line()
```

Learning Assessment

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(tmax_f = (tmax * 9/5) + 32, tmin_f = (tmin * 9/5) + 32) %>% 
  ggplot(aes(x = tmax_f, y = tmin_f)) +
  geom_point(alpha = .4) +
  geom_smooth(method = lm, se = FALSE)
```
?geom_smooth

##Tips and tricks

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 
```

To avoid overplotting---lots of---1000000 data

```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```

```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_bin2d()
```

```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_density2d()
```


Why the two lines below don’t produce the same result:

```{r}
#color setted to blue outside the aes mapping
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin), color = "blue")

#color setted to blue inside the aes mapping, we’re implicitly creating a color variable that has the value blue everywhere; ggplot is then assigning colors according to this variable using the default color scheme.
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin, color = "blue"))
```

## Univariate plots

Histograms

```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()
```

```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()+
  facet_grid(~name)
```

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram()
```

position = "dodge"----places the bars for each group side-by-side
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```


Density plot

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .5)
```

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```


Boxplots
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_boxplot()
```

Violin plot

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin()
```

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  stat_summary(fun.y = median, geom = "point", color = "blue", size = 3)
```

Ridge plots

```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

#Learning Assessment: compare precipitation across locations

highly skewed!

```{r}
ggplot(weather_df, aes(x = prcp)) + 
  geom_density(aes(fill = name), alpha = .5) 
```

```{r}
ggplot(weather_df, aes(x = prcp)) + 
  geom_histogram()
```

```{r}
ggplot(weather_df, aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)
```

```{r}
ggplot(weather_df, aes(y = prcp, x = name)) + 
  geom_boxplot() 
```

compliment this with a plot showing data for all precipitation less than 100, or for a data omitting days with no precipitation
```{r}
weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)

```


##Saving plots

```{r}
weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)
```


